\documentclass[11pt,letterpaper]{article}
\input{../preamble}
\usepackage{amsmath,amsfonts}
\usepackage{hyperref}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{0in}
\setlength{\textheight}{8.2in}
\setlength{\parskip}{0.4em}
\parindent0em
\def\R{\mathbb{R}}
\def\Q{\mathbb{Q}}
\def\N{\mathbb{N}}
\def\Z{\mathbb{Z}}
\def\C{\mathbb{C}}
\begin{document}
\begin{center}
{\bf \Large Math 55a Homework 2}\medskip

Due Wednesday September 17, 2025.
\end{center}

\begin{itemize}
\item You are encouraged to discuss the homework problems with 
other students. However, what you hand in should reflect your 
own understanding of the material. You are NOT allowed to copy solutions 
from other students or other sources. Also, please list at the end of the
problem set the sources you consulted and people you worked with on 
this assignment.
\item Questions marked * may be on the harder side.
\end{itemize}

{\bf Material covered:} Subgroups, normal subgroups, quotients; modular arithmetic,
permutations, etc. (most of Artin chapter 2).
\bigskip

{\bf 1.} Describe a polygon in $\R^2$ whose symmetry group is $\Z/3$. What
is the fewest number of vertices that such a polygon can have?
\medskip

{\bf 2.} Let $G$ be the set of {\em affine transformations} of the real
line, i.e.\ maps $f_{a,b}:\R\to \R$ of the form $f_{a,b}:x\mapsto ax+b$ for
some $a,b\in \R$ with $a\neq 0$.

(a) Show that $G$ is a group, with group law given by composition.

(b) Show that the subsets $H=\{f_{a,b}\,|\,a=1\}$ and $K=\{f_{a,b}\,|\,b=0\}$
are subgroups of $G$. Which one is normal, and what is the quotient of
$G$ by that subgroup?
\medskip

{\bf 3.} Let $G=S^1=\{z\in \C\,|\,|z|=1\}$ be the group of complex numbers
of modulus 1, with multiplication. Let $H\subset G$ be the subgroup $\{\pm 1\}$.
Show that the quotient $G/H$ is isomorphic~to~$G$.\medskip

{\bf 4.} Show that, if $G$ is a finite group and $H\subset G$ is a subgroup
of index 2 (i.e., $|G|/|H|=2$), then $H$ is a normal subgroup of $G$.
\medskip

\begin{proof}[]
  We are given \( |G|/|H|=2 \). Both left and right cosets, then, must give the same partition of \( G \); thus, for the left coset partition either \( g \in  H\implies gH=H \) or \( g \not\in H \implies gH=G \setminus H \), and, for the right coset partition, either \( g \in  H \implies g \not\in H \implies Hg= G \setminus H \); thus both result in \( \{H,G \setminus H\} \implies gH=Hg=G\setminus H , \, \forall  g \not \in  H\). Thus \( gHg^{-1}=(gH)g^{-1}=(Hg)g^{-1}=H \implies H \unlhd G \).
\end{proof}



{\bf 5.} Let $G$ be the group of rotations preserving a cube in $\R^3$.
Show that $G$ is isomorphic to the symmetric group $S_4$ (permutations of a 
four-element set). (Hint: which four-element set?)
\medskip

\begin{proof}[]
Let \( G \) be the octahedral group \( O_h \) for \( \mathbb{R}^3 \). We show that it is a homomorphism: let \( \varphi  \) be a map \( \varphi  : O_h \to S_{4} \). We want to show \( \varphi (ab)=\varphi (a)\varphi (b) , \forall a,b\in O_h\). Consider the group of rotations perserving a cube in \( \mathbb{R}^3 \), notice if we have a cube of length \( 1 \), the coordinates of the verticies are \( (1,1,1),(1,-1,1),(-1,-1,1),(-1,1,1),(1,1,-1),(1,-1,-1),(-1,1,-1),(-1,-1,-1) \); the rotations then across the diagonals that preserve the cube are:

\begin{align*}
d_1 &: (-1,-1,-1) \to (1,1,1) \\
d_2 &: (-1,-1,1) \to (1,1,-1) \\
d_3 &: (-1,1,-1) \to (1,-1,1) \\
d_4 &: (1,-1,-1) \to (-1,1,1).
\end{align*}

Let \( \rho  \) be a rotations in \( O_h \). Then \( \varphi (\rho ) \) is the permutations of the diagonal rotations \(\{d_{1},d_{2},d_{3},d_{4}\} \). Consider two diagonal rotations \( \rho_{1}\) and \( \rho_{2} \in O_h\). 



\begin{align*}
  \varphi (\rho_{1} \rho_{2}) &\stackrel{\mathclap{\normalfont\mbox{?}}}{=} \varphi (\rho_{1}) \varphi (\rho_{2})\\
.\end{align*}


\end{proof}


{\bf 6*.} The dihedral group $D_n$ is the group of order $2n$
consisting of all symmetries (rotations and reflections) of the regular $n$-gon in the plane.
Find all normal subgroups of $D_n$. (Hint: the answer is slightly different
for even $n$ vs.\ for odd $n$.)
\medskip

{\bf 7.} Prove or provide a counterexample: if $A$ is a normal subgroup of
$B$ and $B$ is a normal subgroup of $C$, then $A$ is a normal subgroup of
$C$.
\medskip

\begin{proof}[Proof Other Information]
We will show that \( A\coloneqq  \{ (), (12)(34)\} \) is a normal subgroup of \\ \( B\coloneqq V_4=\{ (),(12)(34),(13)(24),(14)(23)\} \), the normal Klein 4 group. The Klein 4 group will also be shown to be a normal subgroup of \(C\coloneqq  A_{4} \), where 
\[
  A_{4}=\{ (),(123),(132),(234),(243),(341),(314),(412),(421),(12)(34),(13)(34),(14)(23) \}
.\] 
We end will end with the notion that while it is true that \( A\unlhd B \) and \( B\unlhd C \), it is not the case that \( A\unlhd C \). 


  First, \( B \) is a subgroup of \( C \), since it is a subset consisting of all the disjoint cycles in \( C \) closed under the same operation and each element in \( B \) is its own inverse. Since it is closed under both and is a subset, it is obviously a subgroup. An additional note is that \( B \) is an abelian group, shown by the following Cayley table:


  \begin{center}
  \noindent\begin{tabular}[h!]{c | c c c c}
     & 0 & 1 & 2 & 3   \\
    \cline{1-5}
    0 & 0 & 1 & 2 & 3  \\
    1 & 1 & 0 & 3 & 2 \\
    2 & 2 & 3 & 0 & 1 \\
    3 & 3 & 2 & 1 & 0 \\
\end{tabular}
\end{center}



  Second, to show that \( B\unlhd C \), consider a new set, say \( \Pi \) which is the set of possible ways to partition \( \{ 1,2,3,4\} \) into two unordered pairs, so \( \Pi = \{ \pi_{1},\pi_{2}, \pi_{3}\} \), where \( \pi_{1}=\{ \{1,2 \}, \{3,4\} \}, \pi_{2}=\{ \{1,3 \},\{2,4\} \}, \text{ and } \pi_{3}=\{\{1,4\},\{2,3\}\}\). Notice that, the permutations of \( A_{4} \) are permutations from \( S_{4} \), additionally notice that the composition law applies for any \( S_n \), and since \( |\Pi | =3 \) we can keep \( S_3 \) and its composition law in mind. If I take a mapping \( \varphi : A_4 \to S_3 \), since \( A_4, S_{3}, S_4 \) are closed under the same composition, notice that any \( \forall a,b, \varphi (ab)=\varphi (a)\varphi (b) \), thus \( \varphi  \) is homomophic. Given surjectivity, the image of the map is thereby the whole group \( S_3 \); thus the \( \ker (\varphi)\) can be computed, to save you the visual trouble, I have already parsed it, and it should be, by now, clear to see that the only mappings of the permutations on the elements of \( \Pi \) to \( S_{3} \) are \( \ker(\varphi)=\{e,(12)(34),(13)(24),(14)(23) \} \), notice, however that this is the same as \( V_4 \)! Since we have shown \( \ker (\varphi )= V_4\), then, by first isomorphism theorem, \( V_4 \unlhd A_4 \iff B  \unlhd C\). 

  Recall that \( A \subset B \) and \( B \) is abelian. Thus, for any \( bAb^{-1}=A \implies Abb^{-1}=A \implies Ae=A \implies A=A \), and thus \( A \) is a normal subgroup of \( B \). 

  To demonstrate that \( A \ntrianglelefteq C  \) a singular counterexample will suffice. Let \( c=(123)\in  C \) and \( c^{-1}=(132)\in  C \). Take \( a=(12)(34) \in  A \) and compute \( cAc^{-1}=A \):

  \[
    (123)\circ(12)(34)\circ(132)=(14)(23), \, (14)(23) \neq (12)(34)
  .\] 
  Since \( (14)(23)\in B\), but \((14)(23)\not\in A\) we have \( A \ntrianglelefteq C \).


\end{proof}





 

{\bf 8.} Find the order of the group $GL_2(\Z/2)$ of $2\times 2$ matrices
with entries in $\Z/2$ and nonzero determinant (mod~2).
Is $GL_2(\Z/2)$ isomorphic to another group we've encountered?
\medskip

 




\newpage

{\bf 9*.} (a) Let $p$ be a prime number, and suppose $H\subset S_p$ is a
subgroup of the group of permutations of $\{1,2,\dots,p\}$. Show that
if $H$ contains the $p$-cycle $(123\dots p)$ and a transposition (a
permutation that exchanges two elements and fixes all others), then
$H=S_p$. (In other words: no matter how we choose $1\leq i<j\leq p$,
$S_p$ is generated by $(123\dots p)$ and $(ij)$).

(b) Show that the conclusion of part (a) may be false if we don't assume $p$
is prime (even though it is still true that $S_p$ is generated by $(123\dots p)$ and 
a {\em suitably chosen} transposition).
\medskip

{\bf 10*.} Let $G$ be a finitely generated group, and $H\subset G$ a subgroup
of finite index. Show that $H$ is finitely generated.

(Hint: Choose a finite subset $S\subset G$ containing one representative
of each coset of $H$, so every element of $G$ is the
product of an element of $S$ and an element of $H$. Given a word in the generators of $G$ and their
inverses, how do you rewrite it as the product of an element of $S$ and an
element~of~$H$?)
\medskip

{\bf 11*.} (Optional, extra credit): Let $G$ be a non-abelian finite group,
and consider its center $Z(G)=\{a\in G\,|\,ax=xa\ \forall x\in G\}$.

(a) Show that $G/Z(G)$ is not a cyclic group.

(b) Show that at most $5/8^{\mathrm{ths}}$ of the pairs of elements of $G$
commute, i.e.\ the set $C=\{(a,b)\in G\times G\,|\,ab=ba\}$ satisfies
$|C|\leq \frac58 |G|^2$.

(c) Show that this bound is optimal, i.e.\ there exists a non-abelian finite
group for which $|C|=\frac58 |G|^2$.%
\medskip

{\bf 12*.} (Optional, extra credit):  Given a set $S$, let $\mathcal{E}\subset \mathcal{P}(S)$ be such
that (1) $S\in \mathcal{E}$, (2) if $A\in \mathcal{E}$ then $S-A\in
\mathcal{E}$, (3) if $A,B\in \mathcal{E}$ then $A\cup B\in \mathcal{E}$ and 
$A\cap B\in \mathcal{E}$. 

Prove that if $S$ is finite then there is a set
$T$ and a surjective map $f:S\to T$ such that $\mathcal{E}=\{f^{-1}(A),\ 
A\subset T\}$. What happens if $S$ is infinite?
\medskip






{\bf 13.} How long did this assignment take you?  How hard was it?
What resources did you use, and how much help did you need?
(Remember to list the students you collaborated with on this assignment.)
Did you have any prior experience with this material?

Alec del Pino, Tara Westover, Sasha Something

\end{document}
