\documentclass{article}
\input{../preamble}

\begin{document}

\lecture{1}{Introduction}{June 09, 2025}

\section{Introduction}

This is the overview of the \textit{hardware} and \textit{software} of single variable calculus. The structure of the course is as follows: lectures $\longrightarrow $ text \( \longrightarrow  \) notes \( \longrightarrow  \) study guides \( \longrightarrow  \) learning exercises (psets). Calculus may be viewed as high school mathematics with the additional concept of the limit tossed in. Differential calculus is one that focuses on the concept of instantaneous speed. At a particular instance (say, \( P \)) the instantaneous speed seems to be $0$ because we are not moving! We know, however, that we are not magically stationary during this slice of time relative to observers, however. We don't want to know the average speed (i.e., \( \frac{\Delta x}{\Delta y} \)), but rather the instantaneous speed! So what if we brought the two windows of observation (the beginning and the end of the averages (initial and final)), then the average speed will approach the instantaneous speed of point \( P \). 

I wonder if the concept of frames of reference in physics directly relates to the concept of relativity within regards to two frames of reference approaching a singular point; that is to say, did Einstein take inspiration from the fact that relative-ness of the observers eventually approached something which was different from the instantaneous speed that can be observed from afar? This relative difference seems to be crucial in calculus. Perhaps I am waffling, however, and it is nothing of note. 

% Your notes here...

\end{document}
