\documentclass{report}
\input{../preamble}
\course{math55}
\begin{document}

\lecture{9}{Hermitians, \( \CC\)-spectrals, etc.}

\section{Hermitian inner products}

We now wish to real with inner products on complex vector spaces (as opposed to our previous study on \( \mathbb{R} \)). Hermitian inner products are \( \CC \)-linear, i.e., conjugate-linear in one of the inpts. 

\begin{definition}{Hermitian forms}{hermforms}
A \textbf{Hermitian form} on a \( \CC \)-vector space \( V \) is \( H: V\times V \to  \CC\) s.t.
\begin{enumerate}
  \item \( H \) is \textbf{\textit{sesquilinear}}
    \begin{itemize}
      \item \( H(u,v+w)=H(u,v)+H(v,w)\)
      \item \( H( u, \lambda v) =\lambda  H (u,v)\)
      \item however, \( H(\lambda u, v)= \overline{\lambda }  H(u,v) \), where our \( \overline{\lambda }  \) is our complex conjugate (just \( \overline{z}  \) really). 
    \end{itemize}
  \item \( H \) conjugate symmetric \( H(u,v)=\overline{H(v,u)} \implies H(u,v) \in  \mathbb{R} \)
\end{enumerate}
\end{definition}

\begin{definition}{Hermitian inner products}{herminner}
  A hermitian inner product on \( V \) is a positive definite  \(  (H(u,u)\ge 0 \forall u \neq 0) \)
\end{definition}

\dfn{limit}{limiter}


\begin{mybox}{blue}{Remark 1}
  \[
    \varphi _H : V \to  V^*, \, u \mapsto H(u,\cdot )
  \]
  where \( H \) is \( \CC \)-linear \( V \to  \CC \) and \( H \) is linear in the second input. We say \( \varphi  \) is a \textbf{\textit{complex antilinear}} map, and \( (\varphi (\lambda  u ) = \overline{\lambda } \varphi (u) ) \).
\end{mybox}

In the case that \( H \) is positive def. \( \implies \) nondegenerate \( \forall (u \neq 0)\in  V , \varphi _H(u)=H (u,\cdot )\neq 0\). When we are given a subspace \( S \subset  V \), it must be orthogonal, 
\[ S^\perp = \{v \in  V \mid H(v,w)=0 \forall w \in  S\}   \]
s.t. it is a subspace. Of course, as with any finite dimensional conjugate \( \text{dim}V<\infty \), we have \( (S^{\perp})^\perp \). Continuing on, we note that when \( V = S \oplus S^\perp \), \(S \cap S^\perp = \{0\} \), \[
  u \in  S \cap S^\perp \implies H(u,u) \text{ where the first \( u \in  S\), and the second \( u \in  S^\perp \)} =0 \implies u=0,
\] 
so \( \text{dim}S^\perp = \text{dim} V - \text{dim}S \)

\begin{theorem}{Th}{t}
  \( V \) admits an orhtogonal basis in \( \CC\). The proof is the same as in \( \mathbb{R} \). 
\end{theorem}

\begin{definition}{}{}
  An \textbf{orthonormal basis} of \( (V,H) \) is a basis \( \{e_i\}   \) s.t. 
  \begin{displaymath}
    H(e_i,e_i)=\|e_i\|^2 =1, \, H(e_i,e_j)=0 \forall i\neq j, e_i \perp e_j. 
  \end{displaymath}
  
\end{definition}

some other stuff


...
...
So, in matrix form, \( H(z,w)=\overline{z}^{\text{T}}w =z^*=(\overline{z}_{1}\ldots \overline{z}_n) \) as conjugate transpose. 

Consider fourier series as a (sorta-kinda) example of \( V=C^{\infty} (S^1,\CC) \) w/ is infinitely differentiable \( S^1=\mathbb{R}/\mathbb{Z} \to  \CC, (\iff 1)\)-periodic functions \( \mathbb{R} \to  \CC \). \( L^2 \)-inner product, \[
  H(f,g)=\int_{S^1}\overline{f(t)} g(t)dt = \int_0^1 \overline{f(t)} g(t)dt\]
  and 
  \[
    H(f,f) = \int _0^1|f(t)|^2 > 0\,  \text{ if } f\neq 0
  .\] 
  Note that then \[ f_n(t) =e^{2 \pi  i nt} = \cos (2 \pi  nt) + i \sin  (2 \pi  nt), n \in  \mathbb{Z}\] is orthogonal! So, \( H(f_n,f_m)=\delta _{n,m}  \). 


  \begin{definition}{}{}
 We let \( V \) be a complex space, \( H \) a Hermitian inner product, and \( T:V \to  V \) a linear operator.
  \begin{displaymath}
    \text{the adjoint of } T \text{ is } T^* : V \to  V \text{ s.t. } H(T^*v,w)=H(v,Tw) \forall v,w \in  V
  \end{displaymath}

  \[(\iff H(Tv,w)=H(v,T^*w)) .\] 


\end{definition}{}{}
Some immediate concequences:
\begin{itemize}
  \item \( T \) is \textit{self-adjoint} if \( T^*=T \)
  \item \(  T \) is \textit{unitary} if \( H(Tv,Tw)=H(v,w) \)
  \item \( T \) is normal if \( TT^*=T^*T \)
    unitary operators form a group (omg!) \( U(V)=U(V,H) \subset \Aut(V) = \text{GL}(V) \), \( U(n) \subset  \text{GL}(n,\CC) \), e.g., \( U(1)=S^1=\{z \mid  |z| = 1\} \subset  \CC[*]  \). 
\end{itemize}





\begin{mybox}{orange}{Proposition 1}
    in an orthonom. basis \( \mcM (T^*) = \mcM (T)^*=\overline{\mcM(T)^t}  \) so \( H(Tv,w)=(\mcM v )^*w = v^* (\mcM^* w)=H(v,T^*w) \), self-adjoint operators \( \leftrightarrow \) Hermitian matricies 

\end{mybox}


\begin{definition}{Complex spectral theorem}{compspecthr}
  put it here 
\end{definition}

proof goes here

in such a basis \( \mcM (T) = \begin{pmatrix} \lambda_1 & \, & 0 \\ \, & \ddots& \, \\ 0 & \, & \lambda_n \end{pmatrix}  \)


\subsection{some number theory (non-degenerate symmetric bilinear forms)}

this u can go over later via the notes just put it in here

















\end{document}
